Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2016-12-29 Master Thesis/synopsis.tex
Finish proof 5.2 (i)

\begin{proof}
\begin{enumerate}[(i)]
\item Let $A$ be free-by-finite, i.e. $A = A' \oplus F$ for some finite group $A'$ and free group $F$. Since $\ker(f)\subset A$ is finite, it's actually contained in $A'$.

Now let $B$ be free-by-finite and $K = \ker(f)$. Since the injective map $A/K \to B$ induced by $f$ can be restricted to their torsion subgroups, we see that $(A/K)_\text{tor}$ is finite. Every $a\in A_\text{tor}$ can be written as $a = a' + k$ for some $a'\in A,\ k\in K$ such that $na'\in K$ for some $n\in\mathbb{N}$. Since both $(A/K)_\text{tor}$ and $K$ are finite, $A_\text{tor}$ is finite as well. TODO
\item TODO
\item Subgroups of free-by-finite groups are free-by-finite. Now assume $A'$ is finitely generated. We consider $A'$ as a subgroup of $A$. Let $T=A_\text{tor}$. By replacing $A'$, $A$, $A''$ with $A'/(T\cap A')$, $A/T$, $A''/p(T)$ respectively, we can assume that $A$ is free and $A'$ is free of finite rank. Let $B$ be a basis of $A$. There exists a finite subset $C\subset B$ such that $A'\subset \bigoplus_{b\in C}\mathbb{Z}b$. We get an exact sequence:
\[ 0 \longrightarrow \Big(\bigoplus_{b\in C} \mathbb{Z}b\Big)\Big/A' \longrightarrow A/A' \longrightarrow \bigoplus_{b\in B\setminus C}\mathbb{Z}b \longrightarrow 0 \]
The left abelian group is finitely generated, particularly free-by-finite. The right group is free, hence also free-by-finite. By (ii), $A/A'\cong A''$ is free-by-finite as well.
\item It suffices to show this assertion for finite groups and for $A=\mathbb{Z}$. For $A$ finite we have $nA = 0$ for $n = \#A$. Since $\mathbb{Z}$ has no nilpotent elements, the intersection of all prime ideals of $\mathbb{Z}$ is already trivial. \qedhere
\end{enumerate}
\end{proof}

>>>>

\begin{proof}
\begin{enumerate}[(i)]
\item Let $B$ be free-by-finite and $K = \ker(f)$. Considering $A/K$ as a subgroup of $B$, we see that $A/K$ is free-by-finite. Since $(A/K)_\text{tor}$ and $K$ are finite, $K' = \{a\in A\mid na\in K \text{ for some }n\in\mathbb{N}\}$ is finite as well, in particular $A_\text{tor}$ is finite. In fact $K'=A_\text{tor}$, since $K\subset A_\text{tor}$. Now consider the surjection $A\to A/K\to (A/K)/(A/K)_\text{tor}$. This map has kernel $K'$, thus inducing an isomorphism from $A/A_\text{tor}$ to a free abelian group.

\item TODO
Now let $A$ be free-by-finite, i.e. $A = A_\text{tor} \oplus F$ for some free abelian group $F$, and $K=\ker(f)$. Since $K\subset A$ is finite, $K$ is actually contained in $A_\text{tor}$. Therefore, we see that $A/K$ is free by finite as well, and by extension, the image $I$ of $f$. Set $m = \#B/I$ and consider the canonical map $B\stackrel{\cdot m}{\to} I \to I/I_\text{tor}$. This map has kernel $\{b\in B \mid mb \in I_\text{tor}\} = B_\text{tor}$, making $B/B_\text{tor}$ isomorphic to a subgroup of the free abelian group $I/I_\text{tor}$. Now both $I_\text{tor}$ and $B_\text{tor}/I_\text{tor} \subset B/I$ are finite, hence $B_\text{tor}$ is finite as well.
\item Subgroups of free-by-finite groups are free-by-finite. Now assume $A'$ is finitely generated. We consider $A'$ as a subgroup of $A$. Let $T=A_\text{tor}$. By replacing $A'$, $A$, $A''$ with $A'/(T\cap A')$, $A/T$, $A''/p(T)$ respectively, we can assume that $A$ is free and $A'$ is free of finite rank. Let $B$ be a basis of $A$. There exists a finite subset $C\subset B$ such that $A'\subset \bigoplus_{b\in C}\mathbb{Z}b$. We get an exact sequence:
\[ 0 \longrightarrow \Big(\bigoplus_{b\in C} \mathbb{Z}b\Big)\Big/A' \longrightarrow A/A' \longrightarrow \bigoplus_{b\in B\setminus C}\mathbb{Z}b \longrightarrow 0 \]
The left abelian group is finitely generated, particularly free-by-finite. The right group is free, hence also free-by-finite. By (ii), $A/A'\cong A''$ is free-by-finite as well.
\item It suffices to show this assertion for finite groups and for $A=\mathbb{Z}$. For $A$ finite we have $nA = 0$ for $n = \#A$. Since $\mathbb{Z}$ has no nilpotent elements, the intersection of all prime ideals of $\mathbb{Z}$ is already trivial. \qedhere
\end{enumerate}
\end{proof}
